Writing Trash Bin

### 2024-02-26 writingdumpster.qmd
finished word vomit of noise vs signal section

# Moved over notes
- Error is best mapped to fuzziness, location, and colour value; arrangement, size and transparency are an OK second choice; but saturation, hue, orientation and shape are unacceptable and have no intuitive connection to variance [@Maceachren2012]. 
- @Nathonours found that a scatter plot is better than a line plot if you want to convey the correlation between two time series but we cannot be sure if this was influenced by swapping the distribution depicted or by dropping the irrelevant feature of inexchangeability.
- No only do the graphical elements we map our features to matter, but the direction matters too. Graphical elements that are more fuzzy (fuzziness), further from centre (location), lighter (colour value), poorly arranged (arrangement), smaller (size), more transparent (transparency) are perceived to be more uncertain [@Maceachren2012]. 
- This idea also extends to interval estimation, where questions about probability are best answered with gradients and questions about start and end times are easiest to answer with ambiguation [@Gschwandtnei2016].
- Heuristics can work against us just as much as they can work for us. The sine illusion can cause the confidence interval of a smoothed sine curve to seem wider at the peaks than the troughs, causing us to underestimate uncertainty associated with changing values [@Vanderplas2015]
- mentions uncertainty can be variance, precision, accuracy, reliability, or related concepts @Hullman2016
-  @Spiegelhalter2017 defines risk as risk known probabilities and uncertainty as estimated probabilities, therefore uncertainty is what we have when you cannot accurately quantify risk.
- Authors provide little justification for their chosen response models (e.g. absolute accuracy vs relative measures) @Hullman2016 
- Some questions (e.g. what are the chances that the no.6 bus will arrive first) you can elicit the viewers subjective probability distribution @Hullman2016
- The big data paradox shrinks confidence intervals but magnifies bias [@Bradley2021] because the confidence intervals shrink due to the sample size but the imprecise target group creates bias.
- coloured maps for earthquake risk are easily interpretable @Spiegelhalter2017
- Hurricane risk is not only the path, but also the storm surge and wind speed. @Spiegelhalter2017

@@

# Main section outtakes
Visualisation experiments are highly sensitive to the way questions are asked. For example, a study by @Hullman2015 that compared HOPs, error bar, and violin plots found that many participants reported that group A was more likely to be higher than group B, which was not not plausible given that the mean of B was larger than the mean of A. The authors attributed this obvious mistake to the visual representation, but we believe it was more likely caused by the distributions being shown in the order A then B while the question swapped the order by asking "What is P(B>A)?". This flip is minor, but must be done *after* the participants have worked out the probability. Since the primary task is done they forget to complete the secondary task, similar to forgetting your bank card in an ATM. I suspect this was the cause of the mistake as I, when attempting this question myself, made the exact same mistake. I also constalty lead my bank card in ATMs.

@@

# Previous literature review
#####Taxonomy Benefits
- Distinctions at this taxonomy level seem to be important also untangling helps to understand it
  - Failing to communicate the nature of your uncertainty can result in underestimation or overestimation of failure probabilities [@Kiureghian2009]
  - @Gustafson2019 found that the framing of our uncertainty, (i.e. if the source of uncertainty is from a lack of knowledge, approximations, unknown unknowns, or disagreement among parties) was found to not have a detrimental effect on the belief in the estimates, perceived credibility, or behavioural intentions of the decision makers.

>>>>

# Moved over notes
- @Bella2005 found that most participants were ignorant to the fact that error bars are used for both confidence intervals and standard error bars, two wildly different indicators of precision. (**also the paper sherry sent me**)
- Several studies probide evidence that the usability of uncertainty representations can be highly user and task dependent @Kinkeldey2014
- The tasks for participants were most commonly value retrival (some where uncertainty and data value are retrived separately, others where they are some separately). also also be asked to extract a relative value @Kinkeldey2014
- most studies used coincident approaches, static visualisations, are applied to a specific domain and dont have generalised results @Kinkeldey2014
- Most studies involve intrinsic approaches such as colour, transparency, and those that used extrinsic techniques typically used glyphs, error bars, grid based techniques, or contouring @Kinkeldey2014
- Mentioned alleatory vs epistemic uncertainty, in the visualisation literature alleatory uncertainty is more commonly the focus. also mentions dependence and joint pdfs as a consideration @Hullman2016
- When asking questions, researchers should focus on events that can be repeated (because of a misunderstanding of confidence intervals) @Hullman2016 (moved here because it is clear she is touching on the whole "what are we uncertain about" thing)
- we should aim to show enough information to solve a task while avoiding irrelevant distracting information [@kosslyn2006graph]. 
- people will discount information they perceive as discountable if they are able. Including mean estimates on depictions of mass can cause people to discount the uncertainty information and use the difference between means as a proxy for the probability distribution [@Kale2021].
- rarely used extension is aggregation of uncertainty over an area (retrieve overall estimation from a spatial distribution of uncertainty). Other tasks include comparisons, rankings. also have a task called "search" where participants have to identify entities that fulfill certainty characteristics (high or low values) @Kinkeldey2014
- Whether to use classed or unclassed uncertainty is rarely discussed @Kinkeldey2014. e.g. land map with remotely sensed images has ambiguity in the boundaries, which can result from multiple sources (land cover class, measurement error, images from different dates). If you combine them all it might not make sense for someone who only needs one.
- The way we ask questions and the types of questions we ask are selected with little justification. This paper makes suggestions to reduce the noise in the data from these papers. @Hullman2016
- Error is best mapped to fuzziness, location, and colour value; arrangement, size and transparency are an OK second choice; but saturation, hue, orientation and shape are unacceptable and have no intuitive connection to variance [@Maceachren2012]. 
- @Nathonours found that a scatter plot is better than a line plot if you want to convey the correlation between two time series but we cannot be sure if this was influenced by swapping the distribution depicted or by dropping the irrelevant feature of inexchangeability.
- No only do the graphical elements we map our features to matter, but the direction matters too. Graphical elements that are more fuzzy (fuzziness), further from centre (location), lighter (colour value), poorly arranged (arrangement), smaller (size), more transparent (transparency) are perceived to be more uncertain [@Maceachren2012]. 
- This idea also extends to interval estimation, where questions about probability are best answered with gradients and questions about start and end times are easiest to answer with ambiguation [@Gschwandtnei2016].
- Heuristics can work against us just as much as they can work for us. The sine illusion can cause the confidence interval of a smoothed sine curve to seem wider at the peaks than the troughs, causing us to underestimate uncertainty associated with changing values [@Vanderplas2015]
- mentions uncertainty can be variance, precision, accuracy, reliability, or related concepts @Hullman2016
-  @Spiegelhalter2017 defines risk as risk known probabilities and uncertainty as estimated probabilities, therefore uncertainty is what we have when you cannot accurately quantify risk.
- Authors provide little justification for their chosen response models (e.g. absolute accuracy vs relative measures) @Hullman2016 
- Some questions (e.g. what are the chances that the no.6 bus will arrive first) you can elicit the viewers subjective probability distribution @Hullman2016
- The big data paradox shrinks confidence intervals but magnifies bias [@Bradley2021] because the confidence intervals shrink due to the sample size but the imprecise target group creates bias.
- coloured maps for earthquake risk are easily interpretable @Spiegelhalter2017
- Hurricane risk is not only the path, but also the storm surge and wind speed. @Spiegelhalter2017

@@

# Main section outtakes
- tasks e.g. "What is the average of X" (how well lined up you are moving down the data pipeline) and resolution (uncertainty is very often moving back a resolution level in statisitcs - also can be an assumption - bootstrap is a replacement for an assumption, not a resolution change, but bootstrapping is centered on the data)  e.g. "What is the noise around the estimate of the value of X"
Visualisation experiments are highly sensitive to the way questions are asked. For example, a study by @Hullman2015 that compared HOPs, error bar, and violin plots found that many participants reported that group A was more likely to be higher than group B, which was not not plausible given that the mean of B was larger than the mean of A. The authors attributed this obvious mistake to the visual representation, but we believe it was more likely caused by the distributions being shown in the order A then B while the question swapped the order by asking "What is P(B>A)?". This flip is minor, but must be done *after* the participants have worked out the probability. Since the primary task is done they forget to complete the secondary task, similar to forgetting your bank card in an ATM. I suspect this was the cause of the mistake as I, when attempting this question myself, made the exact same mistake. I also constalty lead my bank card in ATMs.

@@

# Previous literature review

##### Bivariate map
Depicts a bivariate map which uses a bivariate colour palette that is created by blending two single hue colour palettes. One colour represents the variable of interest while the other represents the sampling error of that variable. There are two immediate problems with this method. First of all, uncertainty is being expressed with hue and saturation which @Maceachren2012 found to be the worst aesthetics to map to uncertainty to as they don't have an intuitive interpretation. Value has a natural connection to uncertainty (lighter values equate to higher uncertainty and darker values equate to more certainty) so it is a much more appropriate choice. While the `Vizumap` data does depict areas of light and darkness, they are largely irrelevant to the uncertainty measure causing our heuristics to lead us to the incorrect conclusions. The Value-Suppressing Uncertainty Palettes (VSUP) shown in @fig-vsup maps estimates to the hue and error to the value thereby creating a more intuitive plot [@Correll2018]. Additionally, at high levels of uncertainty VSUP only has one output colour, which prevents viewers from decrypting any particular value and also avoids enforcing a binary encoding of significance [@Correll2018]. Unfortunately VSUP are not easy to combine with packages like `Vizumap` which leaves it still somewhat difficult to express this encoding in practice, however the combination of a bivariate map with VSUP has shown to improve decisions in the face of uncertainty [@Correll2018].

#####Taxonomy Benefits
- Distinctions at this taxonomy level seem to be important also untangling helps to understand it
  - Failing to communicate the nature of your uncertainty can result in underestimation or overestimation of failure probabilities [@Kiureghian2009]
  - @Gustafson2019 found that the framing of our uncertainty, (i.e. if the source of uncertainty is from a lack of knowledge, approximations, unknown unknowns, or disagreement among parties) was found to not have a detrimental effect on the belief in the estimates, perceived credibility, or behavioural intentions of the decision makers.
